---
title: "DuckDB Notion Extension - Donut Shop Test"
format: html
---

# Testing DuckDB Notion Extension

This notebook tests the DuckDB Notion extension with the traditional donut shop database setup.

## Setup

Load required libraries and environment variables:

```{r setup}
#| warning: false
#| message: false

# Install packages if needed (run once)
# install.packages("dotenv")
# install.packages("duckdb")

# Load libraries
library(dotenv)
library(duckdb)

# Load environment variables from .env file
load_dot_env("../.env")

# Get credentials
notion_token <- Sys.getenv("NOTION_TOKEN")
products_db <- Sys.getenv("NOTION_PRODUCTS_DB")
orders_db <- Sys.getenv("NOTION_ORDERS_DB")
order_items_db <- Sys.getenv("NOTION_ORDER_ITEMS_DB")
suppliers_db <- Sys.getenv("NOTION_SUPPLIERS_DB")

# Verify token is loaded
if (notion_token == "" || notion_token == "your_actual_token_here") {
  stop("❌ Please set your NOTION_TOKEN in the .env file!")
} else {
  cat("✅ Notion token loaded successfully\n")
}

# Display database IDs (for verification)
cat("\n📊 Database IDs:\n")
cat("Products: ", products_db, "\n")
cat("Orders: ", orders_db, "\n")
cat("Order Items: ", order_items_db, "\n")
cat("Suppliers: ", suppliers_db, "\n")
```

## Connect to DuckDB

```{r connect}
# Create DuckDB connection
con <- dbConnect(duckdb::duckdb())

cat("✅ DuckDB connection established\n")
```

## Install and Load Notion Extension

### How it Works (Like VS Code Extensions)

Just like VS Code extensions create a `.vsix` file, DuckDB extensions create a `.duckdb_extension` file:

1. **Build** the extension → Creates `notion.duckdb_extension`
2. **Load** it in DuckDB → `LOAD './build/release/notion.duckdb_extension'`
3. **Use** it → Query Notion databases!

### Option 1: Load Local Build (For Testing)

```{r load-local}
#| eval: false

# Path to your built extension
# After running: cmake .. && make (see README.md ## Building the Extension Locally)
extension_path <- normalizePath("../build/release/notion.duckdb_extension")

# Check if extension exists
if (file.exists(extension_path)) {
  dbExecute(con, sprintf("LOAD '%s'", extension_path))
  cat("✅ Notion extension loaded from:", extension_path, "\n")
} else {
  stop("❌ Extension not found! Build it first:\n",
       "  cd build/release\n",
       "  cmake -DDUCKDB_PATH=../../../duckdb ../..\n",
       "  make\n")
}
```

### Option 2: From Community (Future)

```{r load-community}
#| eval: false

# This will work once published to DuckDB community extensions
dbExecute(con, "INSTALL notion FROM community")
dbExecute(con, "LOAD notion")
```

### Current Status

```{r status}
# Check if extension is built
ext_path <- "../build/release/notion.duckdb_extension"

if (file.exists(ext_path)) {
  cat("✅ Extension file found!\n")
  cat("📁 Location:", normalizePath(ext_path), "\n")
  cat("💡 Uncomment the 'load-local' chunk above to use it\n")
} else {
  cat("❌ Extension not built yet\n\n")
  cat("📋 Build Instructions:\n")
  cat("1. Clone DuckDB: git clone https://github.com/duckdb/duckdb.git\n")
  cat("2. Build DuckDB: cd duckdb && make && cd ..\n")
  cat("3. Build extension:\n")
  cat("   cd duckdb_notion\n")
  cat("   mkdir -p build/release && cd build/release\n")
  cat("   cmake -DCMAKE_BUILD_TYPE=Release -DDUCKDB_PATH=../../../duckdb ../..\n")
  cat("   make\n")
  cat("\nSee README.md ## Building the Extension Locally for details\n")
}
```

## Test 1: Query Products Database

This is the main test - pull all products from your Notion database.

```{r query-products}
#| eval: false

# Once extension is loaded, this will work:
query <- sprintf("SELECT * FROM read_notion('%s')", products_db)

# Execute query
products <- dbGetQuery(con, query)

# Display results
print(products)

# Summary
cat("\n📊 Products Summary:\n")
cat("Total products: ", nrow(products), "\n")
cat("\nColumns: ", paste(names(products), collapse=", "), "\n")
```

## Expected Output

When the extension works, you should see:

- 6 donut products
- Columns: name, category, price, cost, in_stock, daily_limit, calories
- Products like: Glazed Classic, Chocolate Frosted, Jelly Filled, etc.

## Simulated Test (Remove when extension works)

```{r simulated}
# Simulated data for testing the notebook structure
simulated_products <- data.frame(
  name = c("Glazed Classic", "Chocolate Frosted", "Jelly Filled",
           "Boston Cream", "Pumpkin Spice", "Maple Bacon"),
  category = c("Classic", "Classic", "Filled", "Filled", "Seasonal", "Specialty"),
  price = c(2.50, 2.75, 3.00, 3.25, 3.50, 4.00),
  cost = c(0.75, 0.85, 1.00, 1.20, 1.30, 1.50),
  in_stock = c(TRUE, TRUE, TRUE, TRUE, FALSE, TRUE),
  daily_limit = c(200, 150, 100, 80, 0, 50),
  calories = c(250, 280, 320, 340, 310, 380)
)

print(simulated_products)

# Calculate profit margins
simulated_products$profit <- simulated_products$price - simulated_products$cost
simulated_products$margin_pct <- round(100 * simulated_products$profit / simulated_products$price, 1)

cat("\n💰 Profit Analysis:\n")
print(simulated_products[, c("name", "price", "cost", "profit", "margin_pct")])
```

## Cleanup

```{r cleanup}
# Disconnect from DuckDB
dbDisconnect(con, shutdown=TRUE)
cat("✅ Disconnected from DuckDB\n")
```

## Next Steps

1. ✅ Set up .env file with your Notion token
2. ✅ Verify database IDs are correct
3. 🔄 Build the DuckDB Notion extension (see README.md)
4. 🔄 Load the extension in this notebook
5. 🔄 Run the actual query against Notion
6. 🔄 Verify data matches what's in Notion

---

**Note:** This notebook is ready to test once the DuckDB Notion extension is compiled and available!
